---
title: "STAT 440: Homework 5"
author: "Anurag Anand"
date: "19/04/2021"
output: html_document
---

```{r setup, include=FALSE}

library(readr)
library(sqldf)
library(tibble)
library(jsonlite)
library(dplyr)

```


#### Excercise 1

Importing the baseball dataset and printing the descriptor portion of the data set


```{r Question1, echo=TRUE, warning=FALSE, message=FALSE}
# Question 1
baseball = read_csv("https://uofi.box.com/shared/static/w8tycn7kn21vz88qh0euce2mwg76t0d9.csv")

# Printing the descriptor portion of the dataset
str(baseball)
```

#### Excercise 2

Based on the baseball data set created in Exercise 1, how many players made more than $2 million in the 1987 season? Print a table of all players who make more than $2 million

```{r Question2, echo=TRUE, warning=FALSE, message=FALSE}

# finding out the players’ names, team, position, and salary who made more than 2 million USD
mt_2mil = sqldf("
                select 
                Name,
                Team,
                Position,
                Salary
                
                from baseball
                where Salary > 2000
                
                ")

print(as.tibble(mt_2mil))
```


#### Excercise 3

How many players play each Position? Create a frequency table of only those Positions that have at least 10 players.

```{r Question3a, echo=TRUE, warning=FALSE, message=FALSE}

# Finding out how many players play at each position
player_pos_mt10 = as.tibble(sqldf("

                      select Position,
                             count(distinct Name) as `Player Count`
                             
                             from baseball
                             group by Position
                             having count(distinct Name) >= 10
                             order by `Player Count` desc
                "))

print(player_pos_mt10)

```

Creating a frequency table for all the players positions that have less than 10 players:

```{r Question3b, echo=TRUE, warning=FALSE, message=FALSE}


as.tibble(sqldf("

                      select Position,
                             count(distinct Name) as `Player Count`
                             
                             from baseball
                             group by Position
                             having count(distinct Name) < 10
                             order by `Player Count` desc
                "))
```


All the positions don't look valid. The following table contains the same entries as above, but with a valid position name

| Position | Player Count |
|----------|--------------|
| OF       | 4            |
| 3B       | 5            |
| OF       | 2            |
| DH       | 2            |
| SS       | 1            |
| OF       | 1            |
| CF       | 2            |
| 2B       | 2            |
| 1B       | 2            |

#### Excercise 4

Importing the 1996 NBA Season Data (existing as three separate data sets) using R programming language software and the data URLs stint1, stint2, and stint3 and printing the descriptor portion of each dataset

```{r Question4, echo=TRUE, message=FALSE, warning=FALSE}

# Reading stint data
stint1 = fromJSON("https://uofi.box.com/shared/static/vytgfoo28wf9q4tct61j3sk02pwpsjsn.json")
stint2 = fromJSON("https://uofi.box.com/shared/static/3zu8hqubabw8hxqt14qwjpeginf2orgo.json")
stint3 = fromJSON("https://uofi.box.com/shared/static/cfqaufkt73f8iej2v4pf2bsc4b6ok955.json")

# printing the descriptor portion of each data set
print(str(stint1))
print(str(stint2))
print(str(stint3))

```



#### Excercise 5

Performing various operations on the NBA data set and then listing the players alphabetically. We will be showing only the following columns:  player name, player ID, position, age, ptm

```{r Question5, echo=TRUE, message=FALSE, warning=FALSE}

nba1996_1 = sqldf("select *,
                  (AST *1.0) / (TOV*1.0) as ato,
                  PTS*1.0 / MP*1.0 as ptm,
                  ((ThreeP*1.0) / (ThreePA*1.0))*100  as threeptpercentage
                  from
                  (
                  select * from stint1
                  union 
                  select * from stint2
                  union
                  select * from stint3
                  )
                  group by playerID
                  having stint = max(stint)
                  order by ptm desc, playerID
                  ")

as.tibble(sqldf("
                select 
                Player,
                playerID,
                Pos,
                Age,
                ptm
                from
                nba1996_1
                limit 5
                "))

```


#### Excercise 6

creating the "nba1996_2” such that the data is sorted by rebounds in descending order then player ID alphabetically and showing the following columns only: player name, player ID, position, age, rebounds

```{r Question6, echo=TRUE, warning=FALSE, message=FALSE}

# Removing the variable team and stint 
stint1_n = stint1 %>% select(-Tm, -stint)
stint2_n = stint2 %>% select(-Tm, -stint)
stint3_n = stint3 %>% select(-Tm, -stint)

# Aggregating the data
nba1996_2 = sqldf("
                  select Player,
                         PlayerID,
                         Pos,
                         AVG(Age) as Age,
                         sum(G) as G,
                         sum(GS) as GS,
                         sum(MP) as MP,
                         sum(FG) as FG,
                         sum(FGA) as FGA,
                         sum(ThreeP) as ThreeP,
                         sum(ThreePA) as ThreePA,
                         sum(TwoP) as TwoP,
                         sum(TwoPA) as TwoPA,
                         sum(FT) as FT,
                         sum(FTA) as FTA,
                         sum(ORB) as ORB,
                         sum(DRB) as DRB,
                         sum(AST) as AST,
                         sum(STL) as STL,
                         sum(BLK) as BLK,
                         SUM(TOV) as TOV,
                         SUM(PF) as PF,
                         SUM(PTS) as PTS,
                         sum(ORB) + sum(DRB) as rebounds,
                         ((
                         sum(PTS) + 
                         (sum(ORB) + sum(DRB)) + 
                         sum(AST) + 
                         SUM(STL) + 
                         sum(BLK) -
                         ((sum(FGA) - sum(FG)) + (sum(FTA) - sum(FT))) + sum(TOV))*1.0) / (sum(G)*1.0) as efficiency
                         
                         from 
                          (
                          select * from stint1_n 
                          union
                          select * from stint2_n
                          union
                          select * from stint3_n
                          )
                          group by Player, PlayerID, Pos
                          order by rebounds desc, PlayerID
                  ")

# Printing the result
as.tibble(sqldf("
                select 
                Player,
                playerID,
                Pos,
                Age,
                rebounds
                from
                nba1996_2
                limit 5
                "))


```

#### Excercise 7

Which of the 15 most efficient players in the 1996-1997 season are in the NBA Hall of Fame?

```{r Question7, echo=TRUE, message=FALSE, warning=FALSE}

as.tibble(sqldf("
                select 
                Player,
                playerID,
                efficiency
                from
                nba1996_2
                order by efficiency desc
                limit 15
                "))

```


The following players who were the top 15 most efficient player in the 1996 - 97 season also feature on the __NBA Hall of Fame List__:

* **Karl Malone**
* **Shaquille O'Neal**
* **Grant Hill**
* **Charles Barkley**
* **Hakeem Olajuwon**
* **Patrick Ewing**
* **Michael Jordan**
* **Alonzo Mourning**


#### Excercise 8

Importing the subset of the SBA Business Loans Data and printing the subset of the data 

```{r Question8, echo=TRUE, message=FALSE, warning=FALSE}

sba_bl_data = read_delim("https://uofi.box.com/shared/static/liz915g0j5zeg67of0ykhoa1shwnr2at", 
                          "\t", escape_double = FALSE, 
                         col_types = cols(ApprovalDate = col_date(format = "%d-%b-%y"), 
                                          ChgOffDate = col_date(format = "%d-%b-%y"), 
                                          DisbursementDate = col_date(format = "%d-%b-%y"), 
                                          DisbursementGross = col_number(), 
                                          BalanceGross = col_number(), 
                                          ChgOffPrinGr = col_number(), 
                                          GrAppv = col_number(), 
                                          SBA_Appv = col_number()), 
                          trim_ws = TRUE)

print(str(sba_bl_data))

```

#### Excercise 9

The following operations are being done on the SBA dataset:

* Filteration for the loan apporval date between 1970 and 2013
* Filtering out of the NAICS not corresponding to Agriculture, Forestry, Fishing and Hunting
* Filteration for such loans where gross approved amount is greater than gross disbursement
* Creation of a new variable called Loan Discrepancy
* Subsetting of the dataset by selecting certain columns
* Printing of the first 5 and last 5 observations from the data set

```{r Question9, echo=TRUE, warning=FALSE, message=FALSE}

outsiders = sqldf("
                  select 
                  Name,
                  City,
                  State,
                  BankState,
                  ApprovalFY,
                  NoEmp,
                  abs(GrAppv - DisbursementGross) as loandiscrepancy
                  from sba_bl_data
                  where FLOOR(NAICS/10000) != 11 
                  and GrAppv > DisbursementGross
                  and ApprovalFY >= 1970 and ApprovalFY <= 2013
                  order by ApprovalFY
                  ")

# Printing first 5 obs
print(head(outsiders),5)

# printing last 5 obs
print(tail(outsiders),5)

```


#### Excercise 10

Using the outsider data to create a single dataset that contains only the missing values in the data and their corresponding columns. Then printing the number of unique businesses observations in this dataset

```{r Question10, echo=TRUE, warning=FALSE, message=FALSE}

# Filtering out the businnesses where there is missing value in any of the rows
missing_info_buis = sqldf("
                            select Name
                            from outsiders
                            where Name is null 
                            or City is null
                            or State is null
                            or BankState is null
                            or ApprovalFY is null
                            or NoEmp is null
                            or loandiscrepancy is null
                          
                          ")

print(length(unique(missing_info_buis$Name)))


```

