---
title: "STAT 440: Homework 03"
author: "Anurag Anand"
date: "16/03/2021"
output: html_document
---
```{r include=FALSE}

library(readr)
library(dplyr)
library(stringr)
library(lubridate)

```

### Excercise 1

**Show the distributions of all continuous variables in a visualization. Check the data key for a reasonable range of values among these variables. Flag the points that are outside of the reasonable range of values.**

### Excercise 2

Importing the Urbana Rental Inspections Data using R and formatting the input columns

```{r echo=TRUE}

# Reading the data file and printing the descriptor portion of the data 
urbana_rental_data <- read_csv("https://uofi.box.com/shared/static/b6u4it6rjdjva4ouk4fps7h7ias0u4eu.csv", 
    col_types = cols(`Expiration Date` = col_date(format = "%m/%d/%Y"), 
        `Inspection Date` = col_date(format = "%m/%d/%Y")))

print(str(urbana_rental_data))
```

The shape of the imported data frame is 
```{r}

print(paste("Number of Rows in the data set:",nrow(urbana_rental_data)))
print(paste("Number of Columns in the data set:",ncol(urbana_rental_data)))

```

### Excercise 3

Cleaning the adress column

```{r AddressCleaning, echo=TRUE}
# Extracting the latitude
urbana_rental_data$Latitude = str_extract(urbana_rental_data$`Mappable Address`,"4\\d+\\.\\d+")
urbana_rental_data$Longitutde = str_extract(urbana_rental_data$`Mappable Address`,"\\-\\d+\\.\\d+")

# Writing a function to calculate the adress end
Address_end = function(x){
  str_locate(x, "\\(")[1]
}

# Extracting the index where the adress ends and the latitutde and longitude begins
urbana_rental_data$Address_end_idx = sapply(urbana_rental_data$`Mappable Address`, Address_end)

# Extracting the sub part of the adress before the co-ordinates
urbana_rental_data$New_Address = str_sub(urbana_rental_data$`Mappable Address`,
                                        1,
                                        urbana_rental_data$Address_end_idx-1)

# recombning the co-ordinates in the proper order
urbana_rental_data$New_Address = str_c(urbana_rental_data$New_Address,
                                      "(",urbana_rental_data$Latitude,",",
                                      urbana_rental_data$Longitutde,")")

# Removing all the intermediate columns created and renaming the new adress column
urbana_rental_data = urbana_rental_data %>% select(-c(`Mappable Address`,
                                                     "Latitude",
                                                     "Longitutde",
                                                     "Address_end_idx")) %>% rename(`Mappable Address`=New_Address)

print(head(urbana_rental_data,5), width = Inf)
```

### Excercise 4

Finding out the distance to downtown using the haversine formula and the co-ordinates of central urbana **(40.1106 degrees North, 88.2073 degrees West)**

```{r HaverSine, echo=TRUE}

# Haversine distance function

HaversineDistance = function(x){
  # Central point co-ordinates
  lat1 = 40.1106
  long1 = -88.2073
  
  # Extracting the latitude and longitude from each row of data
  lat2 = x["Latitude"]
  long2 = x["Longitude"]
  
  # Setting up the constants pi and R (Radius of Earth in Km)
  pi = 3.141593
  R = 6371
  
  # Haversine calculation
  phi1 = lat1 * (pi/180)
  phi2 = lat2 * (pi/180)
  del_phi = (lat2 - lat1)*(pi/180)
  del_lambda = (long2 - long1)*(pi/180)
  a = sin(del_phi/2)^2 + (cos(phi1)*cos(phi2)*sin(del_lambda/2)^2)
  c = 2*atan2(sqrt(a),sqrt(1-a))
  
  d = (R*c)/1.609344
  
}

# Creating the distance to downtown calculation
urbana_rental_data = urbana_rental_data %>% mutate(
  Latitude = as.numeric(str_extract(`Mappable Address`,"4\\d+\\.\\d+")),
  Longitude = as.numeric(str_extract(`Mappable Address`,"\\-\\d+\\.\\d+"))
)

urbana_rental_data = urbana_rental_data %>% mutate(distance2downtown=apply(urbana_rental_data[,c("Latitude","Longitude")], 1, HaversineDistance))

print(head(urbana_rental_data,5), width = Inf)
```

### Excercise 5

Printing out the summary statistics of the **distance2downtown** column:

```{r summary, echo=TRUE}


paste("Min:",round(min(urbana_rental_data$distance2downtown,na.rm = TRUE),2),
      "Max:",round(max(urbana_rental_data$distance2downtown,na.rm = TRUE),2),
      "Median:", round(median(urbana_rental_data$distance2downtown,na.rm = TRUE),2),
      "Mean:", round(mean(urbana_rental_data$distance2downtown,na.rm = TRUE),2),
      "Std. Deviation:", round(sd(urbana_rental_data$distance2downtown,na.rm = TRUE),2))


```

### Excercise 6

Importing the Champaign County Sheriff Office & Jail Dataset and formatting the column types 

```{r ccsodata, echo=TRUE}

ccso_data <- read_delim("https://uofi.box.com/shared/static/lbah1aw0cwghyrelzur0tk3akh65tin6.txt", 
    "|", escape_double = FALSE, col_types = cols(booking_date = col_date(format = "%m/%d/%Y"), 
        booking_time = col_time(format = "%H:%M:%S"), 
        released_date = col_date(format = "%m/%d/%Y"), 
        released_time = col_time(format = "%H:%M:%S")), 
    trim_ws = TRUE)

print(paste0("The shape of the ccso data set is: ",
             str_conv(nrow(ccso_data),'UTF8'),
                      " rows and ",
                      str_conv(ncol(ccso_data),'UTF8')," columns"))
```

### Excercise 7

Creating year, day, hour, minute, second column and then validating it with time difference obtained from other columns

```{r DateValidation, echo=TRUE}

# Combining the date and time columns in a single column
ccso_data$booking_date_time = as_datetime(paste(ccso_data$booking_date,ccso_data$booking_time))
ccso_data$release_date_time = as_datetime(paste(ccso_data$released_date,ccso_data$released_time))


# Finding out the time difference in seconds
ccso_data$timediff = with(ccso_data, difftime(release_date_time,booking_date_time,units="secs"))

# Defining the number of seconds in a year, day, hour and minute
year_seconds = 31536000
day_seconds = 86400
hour_seconds = 3600
min_seconds = 60

# Findinng out the jail years
ccso_data$jail_years = as.integer(ccso_data$timediff/year_seconds)

# Finding out the jail days
ccso_data$jail_days = as.integer((ccso_data$timediff - (ccso_data$jail_years*year_seconds))/day_seconds)

#Finding out the jail hours
ccso_data$jail_hours = as.integer((ccso_data$timediff - 
                                     ((ccso_data$jail_years*year_seconds)+
                                        (ccso_data$jail_days*day_seconds)))/hour_seconds)

# Finding out the jail minutes
ccso_data$jail_minutes = as.integer((ccso_data$timediff - 
  (ccso_data$jail_years*year_seconds+
     ccso_data$jail_days*day_seconds+
     ccso_data$jail_hours*hour_seconds))/min_seconds)

# Finding out the jail seconds
ccso_data$jail_seconds = ccso_data$timediff - 
  (ccso_data$jail_years*year_seconds+
     ccso_data$jail_days*day_seconds+
     ccso_data$jail_hours*hour_seconds+
     ccso_data$jail_minutes*min_seconds)


## Creating the two time spent in jail columns
ccso_data = ccso_data %>% mutate(
  time_spent_in_jail1 = (jail_years * 8760) + (jail_days * 24) + jail_hours +(jail_minutes/60) + (jail_seconds/3600)
)

ccso_data$time_spent_in_jail2 = with(ccso_data, difftime(release_date_time,booking_date_time,units="hours"))

# Checking whether the two columns are equal or not by comparing the sums of the two columns

print(sum(ccso_data$time_spent_in_jail1,na.rm = TRUE))
print(sum(ccso_data$time_spent_in_jail2,na.rm = TRUE))

# Printing the first 5
print(head(ccso_data[,c("jacket_number",
                        "booking_date",
                        "booking_time",
                        "released_date",
                        "released_time",
                        "time_spent_in_jail1",
                        "time_spent_in_jail2")],5), width = Inf)

# ... and the last 5 observations
print(tail(ccso_data[,c("jacket_number",
                        "booking_date",
                        "booking_time",
                        "released_date",
                        "released_time",
                        "time_spent_in_jail1",
                        "time_spent_in_jail2")],5), width = Inf)

```

### Excercise 8

Filtering the data for four towns in Champaign County (Champaign, Rantoul, Savoy, and Urbana)

```{r FilteringCity, echo=TRUE}


pattern = "^RAN\\w+L|^RAN\\w+U|^UR\\w+|^UU\\w+|^SAV\\w+|^CHA\\w+GN|^CHA\\w+NG|^CHA\\w+IN|^CHA\\w+\\|"
cc = ccso_data[which(grepl(pattern,ccso_data$city)),]

# writing a function for correcting the city name name
CorrectCityName = function(x){
  if (grepl("^CHA\\w+GN|^CHA\\w+NG|^CHA\\w+IN|^CHA\\w+\\|",x)) {
    "CHAMPAIGN"
  } else if (grepl("^UR\\w+|^UU\\w+",x)){
    "URBANA"
  } else if (grepl("^SAV\\w+",x)){
    "SAVOY"
  } else {
    "RANTOUL"
  }
}

# Looping through all the rows
cc$city_correct = lapply(cc$city, CorrectCityName)

# Sorting the data set on the corrected city name and selecting the given columns

cc_sort = cc %>% arrange(city_correct) %>% select(jacket_number,
                                             `age arrest`,
                                             city,
                                             state,
                                             sex,
                                             race)

# printing the first 5 rows of the data set
print(head(cc_sort,5),width = Inf)

# printing the last 5 rows of the data set
print(tail(cc_sort,5),width = Inf)

```

### Excercise 9

Using the dataset created in the previous excercise finding out the average jail time spent by two groups **(black people or non-black people)**:

```{r JailTime, echo=TRUE}

## Finding out the jail time in hours 
cc$time_spent_in_jail2 = with(cc, difftime(release_date_time,booking_date_time,units="hours"))

## Creating a "Black" vs "Non-Black" Flag
cc$race_flag = sapply(cc$race, function(x){ifelse(x == "Black", "Black","Non-Black")})

## summarizing the data set to get the average time spent in jail by both the categories

cc %>% filter(!is.na(race_flag)) %>% 
  group_by(race_flag) %>% 
  summarise(avg_jail_time = mean(time_spent_in_jail2,na.rm = TRUE))

```

Looking at the table above we can see that the Black people have a higher jail time on average


### Excercise 10

Defining recidivism: If there are multiple booking time against a particular jacket number that indicates a repeated offense. So I would be taking distinct count of booking date time against each jacket number and wherever that is more than 1, it will be an indication of repeated offense

```{r recidivism, echo=TRUE}

## We are first selecting the two concerned columns and then grouping by jacket_number
recidivists = ccso_data %>% 
  select(jacket_number,booking_date_time) %>% 
  group_by(jacket_number) %>% 
  summarise(offense_count = n_distinct(booking_date_time)) %>% filter(offense_count > 1) %>% nrow()

print(recidivists)
```

Therefore the number of repeate offender is **1387**