---
title: "STAT 440: Homework 04"
author: "Anurag Anand"
date: "27/03/2021"
output: html_document
---

```{r setup, include=FALSE}

library(readr)
library(dplyr)
library(jsonlite)

```


### Excercise 1

Based on the image, which of the following best describes how the data is arranged ?

* Sorted by Date then State

### Excercise 2

Importing the NBA Season Data (as three separate data sets)

```{r NbaData, echo=TRUE, message=FALSE, warning=FALSE}

## Reading in three basket ball data sets as separate files and then printing out the 
## first five observations of each data set

stint1 = fromJSON("https://uofi.box.com/shared/static/vytgfoo28wf9q4tct61j3sk02pwpsjsn.json")

stint2 = fromJSON("https://uofi.box.com/shared/static/3zu8hqubabw8hxqt14qwjpeginf2orgo.json")

stint3 = fromJSON("https://uofi.box.com/shared/static/cfqaufkt73f8iej2v4pf2bsc4b6ok955.json")

```


First 5 rows of the `stint1` data set

```{r stint1, echo=TRUE, message=FALSE, warning=FALSE}

head(stint1,5)

```


First 5 rows of the `stint2` data set

```{r stint2, echo=TRUE, message=FALSE, warning=FALSE}

head(stint2,5)

```


First 5 rows of the `stint3` data set

```{r stint3, echo=TRUE, message=FALSE, warning=FALSE}

head(stint3,5)

```

### Excercise 3

Creating a data set that lists the top players with highest points to minutes played ratio

```{r Excercise3, echo=TRUE, message=FALSE, warning=FALSE}

# Combining the three files in 1 file 

nba1996_1 = rbind(stint1,stint2,stint3)

# Filtering the most recent data for each player and creating the columns 
# ato = assists divided by turnovers
# ptm = points divided by minutes played
# threeptpercentage = three-point shots made out of three-point attempts multiplied by 100

# After that sorting the data by ptm (descending), playerid alphabetically, and then selecting only
# the following columns:
# player name, player id, position, age, ptm


nba1996_1 = nba1996_1 %>% 
            group_by(PlayerID) %>% 
            mutate(max_stint = max(stint)) %>% 
            filter(stint == max_stint) %>% select(-max_stint) %>% 
            mutate(ato = AST / TOV,
                   ptm = PTS / MP,
                   threeptpercentage = (ThreeP / ThreePA)*100) %>% 
            arrange(desc(ptm), PlayerID) %>% 
            select(Player, PlayerID, Pos, Age, ptm)

print(head(nba1996_1,5))

```

### Excercise 4
Creating a data set that lists the top players with highest rebounds

```{r Excercise4, echo=TRUE, message=FALSE, warning=FALSE}

nba1996_2 = rbind(stint1,stint2,stint3)

nba1996_2 = nba1996_2 %>% 
            select(-c(Tm, stint)) %>% 
            group_by(Player, PlayerID, Age, Pos) %>% summarise_all(list(sum))


# Creating new variables called rebounds and efficiency
nba1996_2 = nba1996_2 %>% 
            mutate(rebounds = ORB + DRB) %>%
            mutate(efficiency = ((PTS + 
                                rebounds + 
                                AST + 
                                STL + 
                                BLK - 
                                ((FGA - FG) + (FTA - FT)) + 
                                TOV) / G)) %>% arrange(desc(rebounds), PlayerID)
                               

# Listing the top players with highest rebounds
head(nba1996_2 %>% select(Player, PlayerID, Pos, Age, rebounds),5)

```

### Excercise 5

Finding out the list of most efficient players in the __NBA Season (1996 - 97)__ and then comparing it with the NBA Hall of fame list

```{r Excercise5, echo=TRUE, message=FALSE, warning=FALSE}

# Displaying the list of top 15 players with highest efficiency

head((nba1996_2 %>% ungroup() %>% select(Player, efficiency) %>% arrange(desc(efficiency))),15)

```

The following players who were the top 15 most efficient player in the 1996 - 97 season also feature on the __NBA Hall of Fame List__:

* **Karl Malone**
* **Shaquille O'Neal**
* **Grant Hill**
* **Charles Barkley**
* **Hakeem Olajuwon**
* **Patrick Ewing**
* **Michael Jordan**
* **Alonzo Mourning**


### Excercise 6

Creating the total stints and longest variable and printing the resulting data sets

```{r Excercise6, echo=TRUE, message=FALSE, warning=FALSE}

# Creating the combined data set
nba_1996_3 = rbind(stint1,stint2,stint3)

# Creating a data frame that contains player details and the number of stints they had in this season

nba_1996_3_1 = nba_1996_3 %>% 
              group_by(PlayerID) %>% 
              mutate(total.stints = max(stint)) %>% 
              filter(stint == total.stints)

## Creating a separate dataset that contains the longest duration of games played by each player for a particular team

nba_1996_3_2 = nba_1996_3 %>% 
  group_by(PlayerID, Tm) %>% 
  mutate(total.games = sum(G)) %>% 
  group_by(PlayerID) %>% 
  mutate(longest = max(total.games)) %>% 
  filter(total.games == longest) %>% 
  group_by(PlayerID) %>% filter(stint == min(stint))

# Joining the two data sets created above in single data set 
# joining the two data sets created above on player id. Subsetting them both before joining 
nba_1996_3_1 = nba_1996_3_1 %>% select(Player, PlayerID, Pos, total.stints)
nba_1996_3_2 = nba_1996_3_2 %>% select(Player, PlayerID, Pos, longest)


nba_1996_3 = nba_1996_3_1 %>% 
  inner_join(nba_1996_3_2, by = "PlayerID") %>% 
  select(Player.x ,PlayerID, Pos.x, total.stints, longest) %>% 
  arrange(desc(total.stints), PlayerID)

# Renaming the columns
colnames(nba_1996_3) = c("Player","PlayerID","Pos","total.stints","longest")

# displaying the top 10 rows
head(nba_1996_3, 10)

```

### Excercise 7


Reading the three data sets, namely, `ads_demo_big`, `trips_big` and `delivery_stores_big`

```{r Excecise7, echo=TRUE, message=FALSE, warning=FALSE}

# Reading the ads_demo_big data set from the link provided
ads_demo_big = read_csv("https://uofi.box.com/shared/static/buv7xuweoq51zw4obcf2netx4d16295h.csv")

# Reading the trips_big data set from the link provided. The file is tab delimited
trips_big = read_delim("https://uofi.box.com/shared/static/wmsx1tvrrncryf0ue7ldp4szhizdb3gf.txt", 
    "\t", escape_double = FALSE, trim_ws = TRUE)

# Reading the delivery_stores_big data from the link provided. The file is space delimited and double quotes
# are escaped while reading in the file
delivery_stores_big = read_delim("https://uofi.box.com/shared/static/56bzob5ub1pu5im5fgm1125e4xocbwj2",
                                   " ",escape_double = TRUE, trim_ws = TRUE)

```

Printing the structure of the demographics data.

```{r DemoData, echo=TRUE,warning=TRUE, message=FALSE}

print(str(ads_demo_big))

```

Printing the structure of the trips data.

```{r TripsData, echo=TRUE,warning=TRUE, message=FALSE}

print(str(trips_big))

```

Printing the structure of the Store data.

```{r StoreData, echo=TRUE,warning=TRUE, message=FALSE}

print(str(delivery_stores_big))

```

### Excercise 8

Among the panelists with at least one child aged 0-17 years, which panelists spend at least one hundred dollars at least 3 different grocery stores in the first 40 weeks of the year?

```{r Question8, echo=TRUE,warning=TRUE,message=FALSE}

# Subsetting panelist id's for those who have at least one child in the age bracket of 0-17

ads_demo_fil = ads_demo_big %>% 
              filter(`Children Group Code` %in% c(1,2,3,4,5,6,7))


# Filtering the trips data for the first 40 weeks of the year
# week 1: 1114
# week 40: 1114 + (40-1)*1 = 1153

trips_data_filtered = trips_big %>% 
                      filter(WEEK >= 1114 & WEEK <= 1153)

# Joining the filtered demographics data with the filtered trips data

ads_demo_fil %>% 
  inner_join(trips_data_filtered, by = c("Panelist ID"="PANID")) %>% 
  group_by(`Panelist ID`) %>% 
  summarise(Num_Stores = n_distinct(IRI_Key),
            dollars_spent = sum(CENTS999)/100) %>% ungroup() %>% 
            filter(Num_Stores >= 3 & dollars_spent >= 100) %>% arrange(`Panelist ID`) %>% 
            inner_join(ads_demo_fil, by = c("Panelist ID"="Panelist ID")) %>% select(`Panelist ID`,
                                                                                     `Children Group Code`,
                                                                                     dollars_spent,
                                                                                     Num_Stores) %>% head(10)

```


### Excercise 9

Among any panelists who shop at the same grocery stores, how many panelists - who are renting their homes - spend more than five hundred dollars in the first 6 months of the year?
 
```{r Excercise9, echo=TRUE, warning=FALSE, message=FALSE}

## Filtering out the panelists who are renting out their homes

renters = ads_demo_big %>% filter(`Type of Residential Possession` == 1)

## Filtering the trips data for the first 6 months, summarizing it by store and pan id and then joining it
## with the table created above (which contains the panelist id for all the renters)

trips_big %>% filter(WEEK <= 1139) %>% arrange(IRI_Key, PANID) %>% 
            group_by(IRI_Key, PANID) %>% summarise(tot_exp = sum(CENTS999)/100) %>% 
            filter(tot_exp >500) %>% ungroup() %>% inner_join(renters, by = c("PANID"="Panelist ID")) %>% 
            select("PANID",
                   `Type of Residential Possession`,
                   'tot_exp') %>% arrange("PANID") %>% head(10)

```
 
 
### Excercise 10
 
 Which panelists - with more than one person living in their household - spend at least $1000 at 2 or more stores each (if any)?
 
 
```{r Excercise10, echo=TRUE, message=FALSE, warning=FALSE}

# Filtering out for panelists with more than one person living in their household

fam_siz_mt1 = ads_demo_big %>% filter(`Family Size` %in% c(2,3,4,5,6))

# Finding out the expenditure of each panelist at a store level. Then filtering out the stores which have
# expenditure less than 1000 dollars. One we have the list of panelists we find out the number of stores 
# they have had that expenditure at and again filter out the panelists who have done this expenditure at less
# than 2 stores. Finally joining the resultant data set with the data set "fam_siz_mt1"

trips_big %>% 
  group_by(PANID, IRI_Key) %>% 
  summarise(tot_exp = sum(CENTS999)/100) %>% 
  filter(tot_exp >= 1000) %>% group_by(PANID) %>% summarise(num_stores = n_distinct(IRI_Key)) %>% 
  filter(num_stores >=2) %>% inner_join(fam_siz_mt1, by = c("PANID"="Panelist ID")) %>% 
  arrange(PANID) %>% select(PANID,`Family Size`)

```
